---
title: "Data Visualization with ggplot2"
author: "Scott Stoltzman"
date: "5/15/2019"
output: html_document
---

```{r initial_conditions, warnings=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library('tidyverse')
# if there is an error from this, then you need to run the following: install.packages('tidyverse')
```


# Homework - Data Visualization  
-----
#### Tidy Tuesday with ggplot2  


In this assignment you will take data given from the `student_ratio` variable below and make relevant charts from it. Reading the data into R has been taken care of for you, but the visualization is up to you!

In the file `tidy_tuesday.html` you will find how the results of the charts should look.

You are also responsible for completing all "Exercises" in Chapter 1 of R for Data Science. 

```{r import_data, warnings=FALSE, message=FALSE }
# This is the data
data <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")
knitr::kable(head(data))
```


----

### Visualize student_ratio across all countries for all years using a histogram!  

Below, you need to add something after `geom_` in order to create a histogram. What do you need to add?

**Ignore** any warnings for now!

```{r}
# What do you need to add to this to make it work?
ggplot(data, aes(x=student_ratio)) + 
  geom_histogram(binwidth = 0.5)
```

----

### Visualize average Primary Education student_ratio in the USA over the years with a line chart.  

Hint, I have already filtered the data for you, just create a line!


```{r}
filtered_data = data %>% filter(country_code == 'USA', indicator == 'Primary Education')
ggplot(filtered_data, aes(x=year, y=student_ratio)) + 
  geom_point() + 
  geom_line()
```

----

### Visualize student_ratio of each indicator (Primary, Secondary, etc.) in the USA over the years with a line chart. Utilize colors to differentiate between the different indicators.  

Hint, I have already filtered the data for you, just create a line (similar to the one above) and add colors.

```{r}
filtered_data = data %>% filter(country_code == 'USA')
ggplot(filtered_data, aes(x=year, y=student_ratio, color = indicator)) +
  geom_point()+
  geom_line()

```


----

### Visualize a bar plot of the different levels student_ratio by country for Primary Education in 2015 by country (of the top 10 biggest student_ratio)  

Hint, data is filtered for you


```{r}
filtered_data = data %>% 
  filter(indicator == 'Primary Education', year == 2015) %>%
  arrange(desc(student_ratio)) %>% top_n(10, student_ratio)

ggplot(filtered_data, aes(x=country, y=student_ratio)) + 
  geom_bar(stat = "identity")

```

----

Here we made it into pretty colors just for fun

```{r}
filtered_data = data %>% 
  filter(indicator == 'Primary Education', year == 2015) %>%
  arrange(desc(student_ratio)) %>% top_n(10, student_ratio)

ggplot(filtered_data, aes(x=country, y=student_ratio, fill = country)) + 
  geom_bar(stat = "identity", show.legend = FALSE)

```

#### Flip the plot above on its side and sort it from high to low (highest on top)  

Hint, we will just continue to use the `filtered_data` from the previous cell

```{r}

filtered_data = data %>% 
  filter(indicator == 'Primary Education', year == 2015) %>%
  arrange(desc(student_ratio)) %>% top_n(10, student_ratio)

filtered_data2 <- filtered_data[order(filtered_data$student_ratio),]

ggplot(filtered_data2, aes(x = reorder(country, student_ratio), y=student_ratio, fill = country)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  ggtitle("2015 Primary Education Class Size") +
  labs(x="Country",y="Student Ratio")


```


----

### Answer the following questions:  

From "The Good, The Bad, The Ugly" data visualization lecture:  

**Name 2 things that make a "Good Visualization":**
  - Clearly illustrate a point
  - Are tailored to the appropriate audience
    
  
**Name 2 things that make a "Bad Visualization":**
  - Are difficult to interpret
  - Are unintentionally misleading
  - Contain redundant and boring information
  

**Name 2 things that make an "Ugly Visualization":**
  - Are almost impossible to interpret
  - Are filled with completely worthless information
  - Are intentionally created to mislead the audience
  - Are inaccurate

  
**What type of plot is best for visualizing relationships between two variables?**  
  - Scatter plot?


**When would you use a pie chart?**
  - If I really wanted to irritate Scott Stoltzman
  - Or if I was attempting to show the categorical composition of a total population of data

